---
title: "DAG Validation Report"
date: "`r format(Sys.time(), '%B %d, %Y — %H:%M')`"
output:
  pdf_document:
    toc: false
    latex_engine: pdflatex
header-includes:
  - \usepackage{booktabs}
params:
  amat: NULL
  amat_testable: NULL
  ci_summary: NULL
  ci_rejected: NULL
  ci_all: NULL
  alpha: 0.05
  tests: "gcm, pcm"
  interpretation: ""
  layout_mat: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(igraph)
```

## 1. Overview

This report summarises the DAG falsification analysis performed with the
**DAG Validator Agent**.

- **Significance level:** `r params$alpha`
- **CI test types:** `r params$tests`
- **Number of variables:** `r if (!is.null(params$amat)) ncol(params$amat) else "N/A"`
- **Variables:** `r if (!is.null(params$amat)) paste(colnames(params$amat), collapse = ", ") else "N/A"`

\vspace{6pt}

## 2. Hypothesised DAG

```{r dag-plot, fig.width=6, fig.height=6.5, fig.align='center'}
if (!is.null(params$amat)) {
  A    <- params$amat
  vars <- rownames(A)
  g    <- graph_from_adjacency_matrix(A, mode = "directed")

  if (!is.null(params$layout_mat)) {
    lay <- params$layout_mat[vars, , drop = FALSE]
  } else {
    lay <- layout_in_circle(g)
  }
  xlim <- range(lay[, 1]) * 1.2
  ylim <- range(lay[, 2]) * 1.2

  op <- par(mar = c(4, 1, 2, 1))
  on.exit(par(op), add = TRUE)

  plot(g,
       main = "Hypothesised DAG",
       vertex.label.cex = 1.0, vertex.size = 22,
       edge.arrow.size = 0.35, layout = lay,
       rescale = FALSE, xlim = xlim, ylim = ylim)

  # Overlay CI tests as red dotted lines
  A_test <- params$amat_testable
  if (!is.null(A_test)) {
    for (i in seq_len(nrow(A_test))) {
      for (j in seq_len(ncol(A_test))) {
        if (i < j && A_test[i, j] == 1L) {
          v1 <- rownames(A_test)[i]; v2 <- colnames(A_test)[j]
          if (v1 %in% vars && v2 %in% vars) {
            idx1 <- match(v1, vars); idx2 <- match(v2, vars)
            segments(lay[idx1, 1], lay[idx1, 2],
                     lay[idx2, 1], lay[idx2, 2],
                     col = "red", lty = 3, lwd = 2)
          }
        }
      }
    }
  }

  # Legend
  leg_labels <- "DAG edges (directed)"
  leg_col    <- "black"
  leg_lty    <- 1
  leg_lwd    <- 2
  if (!is.null(A_test)) {
    leg_labels <- c(leg_labels, "CI tests to run (undirected)")
    leg_col    <- c(leg_col, "red")
    leg_lty    <- c(leg_lty, 3)
    leg_lwd    <- c(leg_lwd, 2)
  }
  par(xpd = TRUE)
  legend("bottom", legend = leg_labels, col = leg_col,
         lty = leg_lty, lwd = leg_lwd, bg = "white",
         box.col = "grey80", cex = 0.8, horiz = TRUE,
         inset = c(0, -0.12))
} else {
  plot.new(); text(0.5, 0.5, "No DAG defined.", cex = 1.2)
}
```

\vspace{6pt}

## 3. CI Test Summary

```{r summary-table, results='asis'}
if (!is.null(params$ci_summary) && nrow(params$ci_summary) > 0) {
  knitr::kable(params$ci_summary, format = "latex", booktabs = TRUE)
} else {
  cat("No CI tests were run.")
}
```

\vspace{6pt}

## 4. Rejected CIs

```{r rejected-table, results='asis'}
if (!is.null(params$ci_rejected) && nrow(params$ci_rejected) > 0) {
  knitr::kable(params$ci_rejected, format = "latex", booktabs = TRUE)
} else {
  cat("No CI statements were rejected.")
}
```

## 5. DAG with Rejected CIs

```{r dag-rejected-plot, fig.width=6, fig.height=6.5, fig.align='center'}
if (!is.null(params$amat) && !is.null(params$ci_all) && nrow(params$ci_all) > 0) {
  A    <- params$amat
  vars <- rownames(A)
  g    <- graph_from_adjacency_matrix(A, mode = "directed")

  if (!is.null(params$layout_mat)) {
    lay <- params$layout_mat[vars, , drop = FALSE]
  } else {
    lay <- layout_in_circle(g)
  }
  xlim <- range(lay[, 1]) * 1.2
  ylim <- range(lay[, 2]) * 1.2

  op <- par(mar = c(4, 1, 2, 1))
  on.exit(par(op), add = TRUE)

  plot(g,
       main = "CI Test Results",
       vertex.label.cex = 1.0, vertex.size = 22,
       edge.arrow.size = 0.35, layout = lay,
       rescale = FALSE, xlim = xlim, ylim = ylim)

  # Overlay CI tests as red dotted lines
  A_test <- params$amat_testable
  if (!is.null(A_test)) {
    for (i in seq_len(nrow(A_test))) {
      for (j in seq_len(ncol(A_test))) {
        if (i < j && A_test[i, j] == 1L) {
          v1 <- rownames(A_test)[i]; v2 <- colnames(A_test)[j]
          if (v1 %in% vars && v2 %in% vars) {
            idx1 <- match(v1, vars); idx2 <- match(v2, vars)
            segments(lay[idx1, 1], lay[idx1, 2],
                     lay[idx2, 1], lay[idx2, 2],
                     col = "red", lty = 3, lwd = 2)
          }
        }
      }
    }
  }

  # Overlay rejected CIs as solid red
  rejs <- params$ci_all[params$ci_all$rejected, , drop = FALSE]
  has_rejected <- nrow(rejs) > 0
  if (has_rejected) {
    canon <- function(x) gsub("[^a-z0-9]", "", tolower(sub("_(ll|ul)$", "", x)))
    canon_vars <- canon(vars)
    name_map   <- setNames(vars, canon_vars)

    for (k in seq_len(nrow(rejs))) {
      ci_str <- as.character(rejs$CI[k])
      main_str <- trimws(ci_str)
      main_str <- gsub("_\\|\\|_", "\u27C2", main_str)
      main_str <- gsub("\\|\\|",   "\u27C2", main_str)
      main_str <- sub("\\s*\\|.*$", "", main_str)
      parts <- trimws(strsplit(main_str, "\u27C2", fixed = TRUE)[[1]])
      if (length(parts) < 2) next
      Xc <- canon(parts[1]); Yc <- canon(parts[2])
      if (!(Xc %in% names(name_map) && Yc %in% names(name_map))) next
      i <- match(name_map[[Xc]], vars)
      j <- match(name_map[[Yc]], vars)
      if (is.na(i) || is.na(j)) next
      segments(lay[i, 1], lay[i, 2], lay[j, 1], lay[j, 2],
               col = "red", lty = 1, lwd = 3)
    }
  }

  # Legend
  leg_labels <- "DAG edges (directed)"
  leg_col    <- "black"
  leg_lty    <- 1
  leg_lwd    <- 2
  if (!is.null(A_test)) {
    leg_labels <- c(leg_labels, "CI tests (tested)")
    leg_col    <- c(leg_col, "red")
    leg_lty    <- c(leg_lty, 3)
    leg_lwd    <- c(leg_lwd, 2)
  }
  if (has_rejected) {
    leg_labels <- c(leg_labels, "Rejected CIs")
    leg_col    <- c(leg_col, "red")
    leg_lty    <- c(leg_lty, 1)
    leg_lwd    <- c(leg_lwd, 3)
  }
  par(xpd = TRUE)
  legend("bottom", legend = leg_labels, col = leg_col,
         lty = leg_lty, lwd = leg_lwd, bg = "white",
         box.col = "grey80", cex = 0.8, horiz = TRUE,
         inset = c(0, -0.12))
} else {
  plot.new(); text(0.5, 0.5, "No CI test results available.", cex = 1.2)
}
```

\vspace{6pt}

## 6. Interpretation

```{r interpretation, results='asis'}
if (nzchar(params$interpretation)) {
  cat(params$interpretation)
} else {
  cat("No LLM interpretation available.")
}
```
